%
% $Id: manual.tex,v 1.52.2.3 1996/12/04 19:04:45 revow Exp $
%

\documentclass[12pt]{article}
\usepackage{alltt}
\usepackage{float}
\usepackage{fancyheadings}
\usepackage{mystyle}
% \usepackage{html}
% \usepackage{heqn}
\usepackage{epsf}

% \includeonly{appC}

%
% Resize the area for text on a page.
%

\setlength{\topmargin}{-0.2in}
\setlength{\headheight}{0.2in}
\setlength{\headsep}{0.2in}
\setlength{\textheight}{8.7in}
\setlength{\textwidth}{6.5in}
\setlength{\oddsidemargin}{0.0in}
\setlength{\evensidemargin}{0.0in}

%
% Fancy headings with the string ``DELVE'' on one side of the header,
% and the section number and name on the other. The page number is in
% the center of the footer.
%

\lhead[]{}
\rhead[]{}
\chead[]{}

\pagestyle{fancyplain}

%
% The level number of the least significant sectional unit listed in
% the table of contents. 2 = subsections
%

\setcounter{tocdepth}{2}

%
% Let floats take up *a lot* of space on a page.
%

\renewcommand{\topfraction}{0.8}
\renewcommand{\bottomfraction}{0.8}
\renewcommand{\textfraction}{0.2}

%
% \begin{Session} body \end{Session}
%
% An environment for displaying an example of a command session. Text
% is printed in typewriter font with a slightly smaller point size
% than the default point size.  Text formatting is unchanged as in
% ``verbatim'', but everything is indented by the current listing
% indentation. 
%

\newenvironment{Session}%
    {\begin{list}{}{}\item[]\small\begin{minipage}{\linewidth}\begin{alltt}}%
    {\end{alltt}\end{minipage}\end{list}}

%
% Strings that appear many times in the document.
%

\newcommand{\Dinfo}{{\rm\tt Dinfo}}
\newcommand{\OR}{{\em ~or~~}}
\newcommand{\delvepath}{{\rm\tt DELVE\_PATH}}
\newcommand{\delve}{DELVE}
\newcommand{\dinfo}{{\rm\tt dinfo}}
\newcommand{\dls}{{\rm\tt dls}}
\newcommand{\dmore}{{\rm\tt dmore}}
\newcommand{\dcheck}{{\rm\tt dcheck}}
\newcommand{\dgenproto}{{\rm\tt dgenproto}}
\newcommand{\dgenorder}{{\rm\tt dgenorder}}
\newcommand{\file}[2]{{\rm\tt #1}{\it .#2\/}}
\newcommand{\mgendata}{{\rm\tt mgendata}}
\newcommand{\mgendir}{{\rm\tt mgendir}}
\newcommand{\minfo}{{\rm\tt minfo}}
\newcommand{\mloss}{{\rm\tt mloss}}
\newcommand{\mls}{{\rm\tt mls}}
\newcommand{\mmore}{{\rm\tt mmore}}
\newcommand{\mstats}{{\rm\tt mstats}}
\newcommand{\mtable}{{\rm\tt mtable}}
\newcommand{\wwwhome}{http://www.cs.utoronto.ca/\maketilde{}delve/}
\newcommand{\maketilde}{\raisebox{0.4ex}{\scriptsize $\sim$}}

\def\beq{\begin{eqnarray}}
\def\eeq{\end{eqnarray}}
\def\eep{\end{eqnarray}}

\begin{document}

% ----------------------------------------------------------------------
% The title page
% ----------------------------------------------------------------------

\begin{center}
   \vspace*{1in}
   {\huge The DELVE Manual} \\[26pt]

   {\large C.~E.~Rasmussen, R.~M.~Neal, G.~E.~Hinton, D.~van~Camp,} \\[5pt]
   {\large M.~Revow, Z.~Ghahramani, R. Kustra, and R.~Tibshirani}   \\[20pt]

   {\large Version 1.1} \\
   December 1996
\end{center}

\vfill

\begin{list}{}{
\setlength{\leftmargin}{0.25in}\setlength{\rightmargin}{0.25in}}\item[]\em
    This manual describes the preliminary release of the DELVE environment.
    Some features described here have not yet implemented, as noted.
    Support for regression tasks is presently somewhat more developed than 
    that for classification tasks.\vspace{3pt}

    We recommend that you exercise caution when using this version of 
    DELVE for real work, as it is possible that bugs remain in the software.
    We hope that you will send us reports of any problems you encounter,
    as well as any other comments you may have on the software or manual, 
    at the e-mail address below.  Please mention the version number of
    the manual and/or the software with any comments you send.
\end{list}

\vfill\vfill

\begin{centering}

For the latest DELVE news, visit \wwwhome\\[4pt]
Send comments to delve@cs.utoronto.ca

\end{centering}\vspace{8pt}

\noindent This project was supported by grants from the Natural Sciences and
Engineering Research Council of Canada and the Institute for
Robotics and Intelligent Systems.

%
%  page style has to come after ``\maketitle'' because it tries to
%  make the style ``plain''.
%

\pagenumbering{roman}
\thispagestyle{empty}

\clearpage

% ----------------------------------------------------------------------
% The copyright notice
% ----------------------------------------------------------------------

\thispagestyle{plain}

\vspace*{\fill}
\vspace*{-0.8in}
\begin{centering}
Copyright \copyright\ 1995--1996 by The University of Toronto,\\
Toronto, Ontario, Canada.\\[1ex]
All Rights Reserved\\[2ex]
\end{centering}

\noindent
Permission to use, copy, modify, and distribute this software and its
documentation for {\bf non-commercial purposes only} is hereby granted
without fee, provided that the above copyright notice appears in all
copies and that both the copyright notice and this permission notice
appear in supporting documentation, and that the name of The
University of Toronto not be used in advertising or publicity
pertaining to distribution of the software without specific, written
prior permission.  The University of Toronto makes no representations
about the suitability of this software for any purpose.  It is
provided ``as is'' without express or implied warranty.\vspace{7pt}

{\bf\noindent
The University of Toronto disclaims all warranties with regard to this
software, including all implied warranties of merchantability and
fitness.  In no event shall the University of Toronto be liable for
any special, indirect or consequential damages or any damages
whatsoever resulting from loss of use, data or profits, whether in an
action of contract, negligence or other tortious action, arising out
of or in connection with the use or performance of this software.  
}\vspace{0.4in}

\noindent If you publish results obtained using \delve{}, please cite
this manual, and mention the version number of the software that you
used.

\vspace*{\fill}
\clearpage 

% ----------------------------------------------------------------------
% The table of contents and list of figures.
% ----------------------------------------------------------------------

\renewcommand{\contentsname}{\hfill CONTENTS \hfill}
\pagestyle{plain}
\thispagestyle{plain}
\setlength{\parskip}{3.3pt}
\vfill
\tableofcontents
\vfill
\clearpage

% ----------------------------------------------------------------------
% The body of the document.
% ----------------------------------------------------------------------

%
% Change paragraphs so the first line is *not* indented, and each
% paragraph is separated by 8 points. Don't forget to turn off the
% extra vertical space that would normally be added to the start of an
% environment if it starts its own paragraph (\partopsep).
%

\setlength{\parskip}{8pt}	% Extra vertical space between paragraphs.
\setlength{\parindent}{0em}	% Width of paragraph indentation.
\setlength{\partopsep}{0pt}	% Extra vertical space, in addition to 
				% \parskip and \topsep, added when user
				% leaves blank line before environment.
\setlength{\topsep}{0pt}
\raggedbottom

\pagenumbering{arabic}
\setcounter{page}{1}

\pagestyle{fancyplain}

\include{sec1}
\include{sec2}
\include{sec3}
\include{sec4}
\include{sec5}

% For reasons that are not apparent, the following has to go here in order to
% produce a page break after Section 6.

\addtocontents{toc}{\protect\newpage\protect\vspace*{1pt}} 

\include{sec6}
\include{sec7}
\include{sec8}

\appendix

\include{appA}
\include{appB}
\include{appC}
\include{appD}

\end{document}
